Exploring the Design Space of Differentially Private Bayesian Inference under Hellinger Distance

### 2019-01-22 ICML2019/main.tex
pass on section 5

\section{Private Mechanisms}
\label{sec_mechs}
In this section we will describe and analyse a series of mechanisms.
Section \ref{sec_ell1_dim} presents a first mechanism, $\lapmech$,
based on the Laplace primitive where the sensitivity is measured w.r.t. to $\ell_1$ norm.
Section \ref{sec_ell1_hist}, shows that $\lapmech$'s accuracy can be improved
by performing a better analysis on the sensitivity of the inference process.
Section \ref{sec_ehd}, shows an instance of the exponential mechanism, $\expmech$, where the scoring
function is based on the Hellinger distance, we will see how this mechanisms performs
poorly due to the very high global sensitivity of the inference process w.r.t to this metric.
Section \ref{sec_ehdl} introduces, $\lexpmech$, a variation of $\expmech$ which uses a lower bound
on the global sensitivity called local sensitivity.
Unfortunately, despite improving on accuracy $\lexpmech$ is non-private.
Finally, building on $\lexpmech$, Section \ref{sec_ehds} introduces $\hexpmech$:
which, based on the notion of \emph{smooth} sensitivity, achieves $\epsilon$-differential privacy and
has good accuracy in specific cases.
In presenting these mechanisms we will denote by $\inferd{\valpha}{\vtheta}{\dataobs}$ the posterior  $\dirichlet{\valpha'}$
obtained starting from the prior $\dirichlet{\valpha}$, likelihood $\multinomial{|\dataobs|}{\vtheta}$, and observed data $\dataobs$.
Two different upper bounds to the value of
\begin{align}
  \label{eqn:eq1}
  \max_{\substack{\adj{\dataobs}{\dataobs'},\\
      \dirichlet{\valpha}=\inferd{\valpha}{\vtheta}{\dataobs},\\
      \dirichlet{\valpha'}=\inferd{\valpha}{\vtheta}{\dataobs'}}}||\valpha-\valpha'||_1 
\end{align}
will be used to calibrate the noise in the mechanisms based on the Laplacian primitive.

\subsection{Calibrating noise w.r.t. $\ell_1$ norm ($\lapmech$)}
\label{sec_ell1_dim}
We start off by presenting $\lapmech$, in Algorithm \ref{lapmech}, based on the Laplace primitive.
The idea of $\lapmech$ is to add Laplacian noise with scale proportional to $\mathcal{X}$ to
the differences between the respective numeric parameters of the prior
and posterior distribution. 
\begin{algorithm}
  \caption{$\lapmech$}
  \label{lapmech}
  \begin{algorithmic}
  \INPUT $\dataobs\in\mathcal{X}^n$, $\dirichlet{\valpha}$
  \STATE \quad {\bf let} $\dirichlet{\valpha'} = \inferd{\valpha}{\vtheta}{\dataobs}$
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )\in \mathbb{N}^{|\mathcal{X}|}$ 
  \STATE \quad {\bf For} $i = 1 \dots |\mathcal{X}| - 1$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{|\mathcal{X}|}{\epsilon}}$
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_{|\mathcal{X}|}} = \alpha_{|\mathcal{X}|} + \lfloor n - \sum_{i = 1}^{|\mathcal{X}|-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\tilde{\valpha}$
  \end{algorithmic}
\end{algorithm}
\begin{lem}
  Algorithm \ref{lapmech} is $\epsilon$-d.p.
  \begin{proof} By noticing that $\mathcal{X}-1$ is an upper bound to the value denoted by (\ref{eqn:eq1}). \end{proof}
\end{lem}

\subsection{Improving the sensitivity ($\ilapmech$)}
\label{sec_ell1_hist}
By seeing the parameters of the distributions as histograms, we can give a tighter upper bound to the value denoted by (\ref{eqn:eq1}).
Indeed a constant will do. Algorithm $\ilapmech$ in Algorithm \ref{ilapmech} takes advantage of this fact. 
Indeed, when $\mathcal{X}=2$, 1 is an upper bound for the value of (\ref{eqn:eq1}), while when $\mathcal{X}>2$, 2 is.
In Algorithm \ref{ilapmech} the algorithm $\ilapmech$ is presented.
\begin{algorithm}
  \caption{$\ilapmech$}
  \label{ilapmech}
  \begin{algorithmic}
  \INPUT $\dataobs\in\mathcal{X}^n$, $\dirichlet{\valpha}$
  \STATE \quad {\bf let} $\dirichlet{\valpha'} = \inferd{\valpha}{\vtheta}{\dataobs}$
  \STATE \quad ${\bf let}\ k=\left \{
        \begin{array}{lll}          
          1&\text{if}& |\mathcal{X}|=2\\
          2&\text{otherwise}
        \end{array}
      \right .\ $
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )\in \mathbb{N}^{|\mathcal{X}|}$ 
  \STATE \quad {\bf For} $i = 1 \dots |\mathcal{X}|-1$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{k}{\epsilon}}$
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_{|\mathcal{X}|}} = \alpha_{|\mathcal{X}|} + \lfloor n - \sum_{i = 1}^{|\mathcal{X}|-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\tilde{\valpha}$
  \end{algorithmic}
\end{algorithm}
\begin{lem}
  Algorithm \ref{ilapmech} is $\epsilon$-d.p.
  \begin{proof} By noticing that 1 and 2 are upper bounds to the value denoted by (\ref{eqn:eq1}), respectively when $|\mathcal{X}|=2$ and $|\mathcal{X}|>2$.
  \end{proof}
\end{lem}

\subsection{A new family of mechanisms}
\label{sec_exp}
In this section, we explore  mechanism which we denote by $\expmech$. $\expmech$ is an instance
of the exponential mechanism \cite{talwar} where the scoring function is based on a metric over
probability distributions.
The idea is that $\expmech$ will sample and output a distribution with higher probability when
it is closer to the real posterior; importantly,  \emph{closer} is measured using a metric over probability
distributions as opposed to a metric over the parameters that represent it. In this work we will
use the Hellinger distance, but this approach could be extended to other $f$-divergences. We will
start by describing how the mechanism works in the case of the Beta-Binomial model noticing that
the extension to the Dirichelet-Multinomial model is straightforward.
Given a prior distribution $\betad(\alpha, \beta)$ and an observed data set: $\dataobs\in\{0,1\}^n$,
we define the follwing set  where the mechanism will sample from:
\[
  \candidateset=\{\bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta),\dataobs)\mid \dataobs \in \{0,1\}^n\}.
\]
Where $\bysinfer(\binomial{n}{\theta}, \bprior,\dataobs)$ denotes the posterior beta distribution inferred starting from the parameters in input.
Notice that the above set reduces to:
\[
  \{\betad(\alpha',\beta')\mid \alpha'=\alpha + \Delta\alpha, \beta'=\beta + n-\Delta\alpha\}
\]
We use $\bysinfer(\dataobs)$ to denote the Bayesian inference process on the observed data $\dataobs$
when the other two parameters, that is prior and likelihood, are fixed.

\subsubsection{A first approach: the $\expmech$ mechanism}
\label{sec_ehd}
$\expmech(x,u,\candidateset)$  outputs an element $r\in\candidateset$ with  probability: % proportional to $\exp(\frac{\epsilon u(x,r)}{2 GS})$.
\[
%\underset{z \thicksim \expmech(x,u,\candidateset) }{\Pr}[z=r] = \frac
\frac
  {\exp(\frac{\epsilon\cdot \ux{r}}{2 GS})}
{\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot \ux{r'}}{2\cdot GS})},
\]
where $u(x,r)$ is the Hellinger scoring function: $-\hellinger(\bysinfer(\dataobs), r)$, and $GS$ is the global sensitivity of the scoring function, defined as:
\begin{equation*}
GS = 
\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}}\max_{\{r\in \mathcal{R}\}}
|\hellinger(\bysinfer(\dataobs), r) - \hellinger(\bysinfer(\dataobs'), r)|
\end{equation*}
%
  \begin{algorithm}
  \caption{$\expmech$ in Beta-binomial model}
  \label{mech:expmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$. $GS$ be the global sensitivity of scoring function.  
  \STATE \quad {\bf Initialize} a vector $Pr = (0, \dots, 0 )$ of $|\candidateset|$ length.
  \STATE \quad {\bf For} $r$ in $\candidateset$:
  \STATE \quad \quad calculate the outputting probability of $r$.
  \STATE \quad \quad $Pr[r] = 
  {\exp(\frac{\epsilon\cdot \ux{r}}{2 GS})}
/ {\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot u(\dataobs, r')}{2\cdot GS})}$
\STATE \quad sample $z$ from $\candidateset$ by probability distribution $Pr$.
  \STATE {\bf return} $z$
  \end{algorithmic}
  \end{algorithm}

Lemma \ref{??} gives us a finite bound on $GS$, while lemma \ref{lem_exp_private} proves privacy $\epsilon$ differential privacy of $\expmech$.

\subsubsection{A second approach based on local sensitivity: $\lexpmech$}
\label{sec_ehdl}
Figure \ref{fig_theory} show that  the accuracy of $\expmech$ is poor if compared with the one of $\ilapmech$. This is due to the fact that
$GS$ is relatively big and hence the variance of the distribution $\expmech$ ends up also being too big. In order to solve this issue we
will proceed in two steps. We will first define $\lexpmech$, which instead of using the global sensitivity, will use the local sensitivity to calibrate
the noise. This is known to be a non private approach, but it will guide us towards new insights.
We will then proceed to use a smooth bound the local sensitivity to
calibrate the noise.

  \begin{algorithm}
  \caption{$\lexpmech$ in Beta-binomial model}
  \label{mech:expmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$. $LS$ be the local sensitivity of scoring function.  
  \STATE \quad {\bf Initialize} a vector $Pr = (0, \dots, 0 )$ of $|\candidateset|$ length.
  \STATE \quad {\bf For} $r$ in $\candidateset$:
  \STATE \quad \quad calculate the outputting probability of $r$.
  \STATE \quad \quad $Pr[r] = 
  {\exp(\frac{\epsilon\cdot \ux{r}}{2 LS})}
/ {\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot \ux{r'}}{2\cdot LS})}$
\STATE \quad sample $z$ from $\candidateset$ by probability distribution $Pr$.
  \STATE {\bf return} $z$
  \end{algorithmic}
  \end{algorithm}

$\lexpmech$ shares candidate set and utility function with $\expmech$,
and outputs a result $r \in \mathcal{R}$ with probability:
\[
%\underset{z \thicksim \lexpmech(x,u,\candidateset)}{\Pr}[z=r] = \frac
\frac
  {\exp(\frac{\epsilon \cdot \ux{r}}{2\cdot LS(\dataobs)})}
{\sum_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot \ux{r'}}{2\cdot LS(\dataobs)})},
\]
where $LS(\dataobs)$ is the local sensitivity of the scoring function, defined as:
\begin{equation}\label{eq:local_sensitivity_scoring}
LS(\dataobs)=\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}, r\in \mathcal{R}}\lvert \hellinger(\bysinfer(\dataobs'), r) - \hellinger(\bysinfer(\dataobs'), r)\rvert 
\end{equation}
The above can be proven to simplify to:
\begin{lem}
\label{lem_loc_opt}
\[
LS(\dataobs) = \max\limits_{\dataobs' \in \datauni^n:\adj{\dataobs}{\dataobs'}} \hellinger(\bysinfer(\dataobs), \bysinfer(\dataobs')).
\]
\end{lem}
The interested reader can check the proof in the Appendix.

\subsubsection{A third approach based on smooth sensitivity: $\hexpmech$}
\label{sec_ehds}
In this section, we explore a new mechanism in Exponential family, which can achieve good accuracy and $\epsilon$-differential privacy simultaneously,
by scaling the noise to a smooth upper bound on the local sensitivity.
We first define the smooth sensitivity on a input $\dataobs$. The following definition is specific to binary data, that is it will
be useful to privately release a posterior Beta distribution. The extension to categorical data is straightforward.
\begin{definition}
\label{def_gamma_smooth}
Given a data set $\dataobs$ of size $n$, we define the $\gamma$-smooth sensitivity of $\hellinger(\bysinfer(\dataobs),\cdot)$ as:
\begin{equation}
  \label{eq_smooth}
   S(\dataobs)=\max_{\dataobs' \in \{0,1\}^{n}}\bigg \{ \frac{1}{\frac{1}{LS(\dataobs')} +\gamma \cdot d(\dataobs,\dataobs')} \bigg \},
\end{equation}
where $d$ is the Hamming distance between two data sets.
\end{definition}

\begin{thm}
  \label{thm_gamma_smooth}
  For any $\dataobs, \dataobs'\in\{0,1\}^n$, and $\gamma \in \mathbb{R}^{+}$
  if  $\adj{\dataobs}{\dataobs'}$ then $\frac{1}{S(\dataobs)} - \frac{1}{S(\dataobs')} \leq \gamma$.
\end{thm}
The interested reader will find the proof in Appendix.
We are now ready to define the mechanism $\hexpmech$. $\hexpmech(\dataobs)$ outputs a candidate $r \in \candidateset$ with probability
\[
  \frac {\exp\big(\frac{-\epsilon\cdot \ux{r}}{4\cdot S(\dataobs)}\big)}
  {\sum_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot \ux{r'}}{4\cdot S(\dataobs)})},
\]
where $S(\dataobs)$ is the 1-smooth sensitivity of  $\hellinger(\bysinfer(\dataobs),\cdot)$. {\color{red}{why do we choose 1???}}
  \begin{algorithm}
  \caption{$\hexpmech$ in Beta-binomial model}
  \label{mech:expmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$. $S(\dataobs)$ be the vector of $\gamma-$sensitivity of scoring function. 
  \STATE \quad {\bf Initialize} a vector $Pr = (0, \dots, 0 )$ of $|\candidateset|$ length.
  \STATE \quad {\bf For} $r$ in $\candidateset$:
  \STATE \quad \quad calculate the outputting probability of $r$.
  \STATE \quad \quad $Pr[r] = 
  {\exp(\frac{\epsilon \cdot \ux{r}}{4 S(\dataobs)})}
/ {\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon \cdot \ux{r'}}{4 \cdot S(\dataobs)})}$
\STATE \quad sample $z$ from $\candidateset$ by probability distribution $Pr$.
  \STATE {\bf return} $z$
  \end{algorithmic}
  \end{algorithm}

As already mentioned $\expmech, \lexpmech, \hexpmech$ extends straightforwardly to the Dirichlet-multinomial system.
Indeed, the Hellinger distance between two Dirichlet distribution, unfolds in:
\[
  \hellinger(\dirichlet{\valpha_1}, \dirichlet{\valpha_2}) = \sqrt{1 - \frac{\betaf(\frac{\valpha_1 + \valpha_2}{2})}{\sqrt{\betaf(\valpha_1) \betaf(\valpha_2)}}},
\]
and the $\candidateset$ will consist of the set of all possible posterior Dirichlet
distributions candidates. Also, the smooth sensitivity $S(\dataobs)$
in (\ref{thm_gamma_smooth}) will be computed by letting $\dataobs'$ range
over all the elements in $\datauni^{n}$ adjacent to $\dataobs$. Notice
that $\candidateset$ has $\binom{n + 1}{m - 1}$ elements in this case. {\color{red}{what is m here? and was $\datauni$} defined?}

\section{Privacy Analysis}

\subsection{Privacy of Laplace Mechanism Family}
The following lemmas prove the privacy of all the mechanisms (besides $\lexpmech$) presented so far.
Mechanisms $\lapmech$ and $\ilapmech$ are trivially $\epsilon$-differentially private since the mechanisms
add Laplacian noise proportional to the $\ell_1$-norm sensitivity to the arguments of the posterior distributions.
\begin{lem}
  $\lapmech$, $\ilapmech$ are $\epsilon$-differentially private.
\end{lem}

\subsection{Privacy of Exponential Mechanism Family}
\subsubsection{$\expmech$ $\epsilon-$Differential Privacy}
\begin{lem}
\label{lem_exp_private}
$\expmech$ is $\epsilon-$differentially private by \cite{dwork2014algorithmic}.
\end{lem}

>>>>

\section{Private Mechanisms}
\label{sec_mechs}
In this section we will describe and analyse a series of mechanisms.
Section \ref{sec_ell1_dim} presents a first mechanism, $\lapmech$,
based on the Laplace primitive where the sensitivity is measured w.r.t. to $\ell_1$ norm.
Section \ref{sec_ell1_hist}, shows that $\lapmech$'s accuracy can be improved
by performing a better analysis on the sensitivity of the inference process.
Section \ref{sec_ehd}, shows an instance of the exponential mechanism, $\expmech$, where the scoring
function is based on the Hellinger distance, we will see how this mechanisms performs
poorly due to the very high global sensitivity of the inference process w.r.t to this metric.
Section \ref{sec_ehdl} introduces, $\lexpmech$, a variation of $\expmech$ which uses a lower bound
on the global sensitivity called local sensitivity.
Unfortunately, despite improving on accuracy $\lexpmech$ is non-private.
Finally, building on $\lexpmech$, Section \ref{sec_ehds} introduces $\hexpmech$:
which, based on the notion of \emph{smooth} sensitivity, achieves $\epsilon$-differential privacy and
has good accuracy in specific cases.
In presenting these mechanisms we will denote by $\inferd{\valpha}{\vtheta}{\dataobs}$ the posterior  $\dirichlet{\valpha'}$
obtained starting from the prior $\dirichlet{\valpha}$, likelihood $\multinomial{|\dataobs|}{\vtheta}$, and observed data $\dataobs$.
To minimize notation we will identify the distribution $\dirichlet{\valpha}$ with its vector
of parameters.
Two different upper bounds to the value of
\begin{align}
  \label{eqn:eq1}
  \max_{\substack{\adj{\dataobs}{\dataobs'},\\
      \valpha=\inferd{\valpha}{\vtheta}{\dataobs},\\
      \valpha'=\inferd{\valpha}{\vtheta}{\dataobs'}}}||\valpha-\valpha'||_1 
\end{align}
will be used to calibrate the noise in the mechanisms based on the Laplacian primitive.
Finally, with $\bysinfer{\dataobs}$, and $\bys{\dataobs}{\vtheta}$ we will denote $\inferd{\valpha}{\vtheta}{\dataobs}$ when $\valpha, \vtheta$ or just $\valpha$
are known from the context.

\subsection{Calibrating noise w.r.t. $\ell_1$ norm ($\lapmech$)}
\label{sec_ell1_dim}
We start off by presenting $\lapmech$, in Algorithm \ref{lapmech}, based on the Laplace primitive.
The idea of $\lapmech$ is to add Laplacian noise with scale proportional to $\mathcal{X}$ to
the differences between the respective numeric parameters of the prior
and posterior distribution. 
\begin{algorithm}
  \caption{$\lapmech$}
  \label{lapmech}
  \begin{algorithmic}
  \INPUT $\dataobs\in\mathcal{X}^n$, $\dirichlet{\valpha}$
  \STATE \quad {\bf let} $\valpha' = \inferd{\valpha}{\vtheta}{\dataobs}$
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )\in \mathbb{N}^{|\mathcal{X}|}$ 
  \STATE \quad {\bf For} $i = 1 \dots |\mathcal{X}| - 1$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{|\mathcal{X}|}{\epsilon}}$
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_{|\mathcal{X}|}} = \alpha_{|\mathcal{X}|} + \lfloor n - \sum_{i = 1}^{|\mathcal{X}|-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\tilde{\valpha}$
  \end{algorithmic}
\end{algorithm}
\begin{lem}
  Algorithm \ref{lapmech} is $\epsilon$-d.p.
  \begin{proof} By noticing that $\mathcal{X}-1$ is an upper bound to the value denoted by (\ref{eqn:eq1}). \end{proof}
\end{lem}

\subsection{Improving the sensitivity ($\ilapmech$)}
\label{sec_ell1_hist}
By seeing the parameters of the distributions as histograms, we can give a tighter upper bound to the value denoted by (\ref{eqn:eq1}).
Indeed a constant will do. Algorithm $\ilapmech$ in Algorithm \ref{ilapmech} takes advantage of this fact. 
Indeed, when $\mathcal{X}=2$, 1 is an upper bound for the value of (\ref{eqn:eq1}), while when $\mathcal{X}>2$, 2 is.
In Algorithm \ref{ilapmech} the algorithm $\ilapmech$ is presented.
\begin{algorithm}
  \caption{$\ilapmech$}
  \label{ilapmech}
  \begin{algorithmic}
  \INPUT $\dataobs\in\mathcal{X}^n$, $\dirichlet{\valpha}$
  \STATE \quad {\bf let} $\valpha' = \inferd{\valpha}{\vtheta}{\dataobs}$
  \STATE \quad ${\bf let}\ k=\left \{
        \begin{array}{lll}          
          1&\text{if}& |\mathcal{X}|=2\\
          2&\text{otherwise}
        \end{array}
      \right .\ $
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )\in \mathbb{N}^{|\mathcal{X}|}$ 
  \STATE \quad {\bf For} $i = 1 \dots |\mathcal{X}|-1$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{k}{\epsilon}}$
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_{|\mathcal{X}|}} = \alpha_{|\mathcal{X}|} + \lfloor n - \sum_{i = 1}^{|\mathcal{X}|-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\tilde{\valpha}$
  \end{algorithmic}
\end{algorithm}
\begin{lem}
  Algorithm \ref{ilapmech} is $\epsilon$-d.p.
  \begin{proof} By noticing that 1 and 2 are upper bounds to the value denoted by (\ref{eqn:eq1}), respectively when $|\mathcal{X}|=2$ and $|\mathcal{X}|>2$.
  \end{proof}
\end{lem}

\subsection{A new family of mechanisms}
\label{sec_exp}
In this section, we explore  mechanism which we denote by $\expmech$. $\expmech$ is an instance
of the exponential mechanism \cite{talwar} where the scoring function is based on a metric over
probability distributions.
The idea is that $\expmech$ will sample and output a distribution with higher probability when
it is closer to the real posterior; importantly,  \emph{closer} is measured using a metric over probability
distributions as opposed to a metric over the parameters that represent it. In this work we will
use the Hellinger distance, but this approach could be extended to other $f$-divergences. 
Given a prior distribution $\dirichlet{\valpha}$,
we define the follwing set  where the mechanism will sample from:
\[
  \candidateset=\{\inferd{\valpha}{\vtheta}{\dataobs}\mid \dataobs \in \mathcal{X}^n\}.
\]
In general $|\candidateset|=\binom{n + 1}{|\mathcal{X}| - 1}$, and in the specific case of $|\mathcal{X}|=2$, that is the Beta-Binomial model,
the above reduces to:
\[
  \{(\alpha',\beta')\mid \alpha'=\alpha + \Delta\alpha, \beta'=\beta + n-\Delta\alpha,  \Delta\alpha=\sum x_i, \dataobs\in \{0,1\}^n\}
\]
where $\valpha=(\alpha,\beta)$.

\noindent \textbf{A first approach: the $\expmech$ mechanism.} $\expmech$'s pseudocode is shown in Algorithm \ref{mech:expmech1}.
Given the input $\dataobs\in\{0,1\}^n$ and a prior $\betad(\alpha, \beta)$, the mechanism $\expmech$ outputs an element $r\in\candidateset$ with  probability: 
\[
\frac
  {\exp(\frac{\epsilon\cdot \ux{r}}{2\cdot GS})}
{\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot \ux{r'}}{2\cdot GS})},
\]
where $u(\dataobs,r)$ is the Hellinger based score function:
$-\hellinger(\bysinfer{\dataobs}, r)$, and $GS$ is the global sensitivity of the scoring function, defined as:
\[
GS = 
\max_{\dataobs,\dataobs': \\ \adj{\dataobs}{\dataobs'}}\max_{\{r\in \mathcal{R}\}}
|\hellinger(\bysinfer{\dataobs}, r) - \hellinger(\bysinfer{\dataobs'}, r)|
\]
% 
  \begin{algorithm}
  \caption{$\expmech$}
  \label{mech:expmech1}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, prior: $\dirichlet{\valpha}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'}= \bys{\dataobs}{\valpha}$.   
  \STATE \quad {\bf Initialize} associative map $pr$ of size $|\candidateset|$.
  \STATE \quad {\bf For} $r$ in $\candidateset$:
  \STATE \quad \quad $pr[r] = 
  {\exp(\frac{\epsilon\cdot \ux{r}}{2\cdot GS})}
/ {\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot u(\dataobs, r')}{2\cdot GS})}$
\STATE \quad {\bf sample} in $z$ from probability distribution $pr$.
\STATE {\bf return} $z$
  \end{algorithmic}
  \end{algorithm}
\begin{lem}
  Algorithm \ref{mech:expmech1} is $\epsilon$-d.p.
\end{lem}

\noindent \textbf{A second approach based on local sensitivity: $\lexpmech$.} Figure \ref{fig_theory} shows that  the accuracy of
$\expmech$ is poor if compared with the one of $\ilapmech$. This is due to the fact that
$GS$ is relatively large and hence the variance of the distribution $\expmech$ ends up also being too large. In order to solve this issue we
will proceed in two steps. We will first define, in Algorithm   \ref{mech:lexpmech}, $\lexpmech$, which instead of using the global sensitivity, will use the local sensitivity to calibrate
the noise. This is known to be a non private approach, but it will guide us towards new insights.
We will then proceed to use a smooth bound the local sensitivity to
calibrate the noise.

  \begin{algorithm}
  \caption{$\lexpmech$}
  \label{mech:lexpmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, prior: $\dirichlet{\valpha}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'}= \bys{\dataobs}{\valpha}$.   
  \STATE \quad {\bf Initialize} associative map $pr$ of size $|\candidateset|$.
  \STATE \quad {\bf For} $r$ in $\candidateset$:
  \STATE \quad \quad $pr[r] = 
  {\exp(\frac{\epsilon\cdot \ux{r}}{2\cdot LS(\dataobs)})}
/ {\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot u(\dataobs, r')}{2\cdot LS(\dataobs)})}$
\STATE \quad {\bf sample} in $z$ from probability distribution $pr$.
\STATE {\bf return} $z$
  \end{algorithmic}
  \end{algorithm}

Given $\dataobs, \lexpmech$ samples an element $r \in \mathcal{R}$ with probability:
\[
\frac
  {\exp(\frac{\epsilon \cdot \ux{r}}{2\cdot LS(\dataobs)})}
{\sum_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot \ux{r'}}{2\cdot LS(\dataobs)})},
\]
where $LS(\dataobs)$ is the local sensitivity of the scoring function, defined as:
\begin{equation}\label{eq:local_sensitivity_scoring}
LS(\dataobs)=\max_{\dataobs'\in\mathcal{X}^n,\adj{\dataobs}{\dataobs'}, r\in \mathcal{R}}\lvert \hellinger(\bysinfer{\dataobs'}, r) - \hellinger(\bysinfer{\dataobs'}, r)\rvert 
\end{equation}
Lemma \ref{lem_loc_opt}  provides a useful characterization of the above function, the interested reader can check the proof in the Appendix.
\begin{lem}
\label{lem_loc_opt}
For $\dataobs\in\mathcal{X}^n$, 
\[LS(\dataobs) = \max\limits_{\dataobs' \in \datauni^n:\adj{\dataobs}{\dataobs'}} \hellinger(\bysinfer{\dataobs}, \bysinfer{\dataobs'})\]
\end{lem}

\subsubsection{A third approach based on smooth sensitivity: $\hexpmech$}
\label{sec_ehds}
In this section, we explore a new mechanism, which compromises between $\expmech$ and $\lexpmech$ and  achieves good accuracy and $\epsilon$-differential privacy simultaneously,
by scaling the noise to a smooth upper bound on the local sensitivity.
We first define the smooth sensitivity on a input $\dataobs\in\{0,1\}^n$.
% The following definition is specific to binary data, hence it will
% be useful to privately release only posterior Beta distributions in the Beta-Binomial model. The extension to categorical data is straightforward.
\begin{definition}
\label{def_gamma_smooth}
For $\dataobs\in\mathcal{X}^n$, and $\gamma \in \mathbb{R}^{> 0}$, the $\gamma$-smooth sensitivity of $\hellinger(\bysinfer{\dataobs},\cdot)$ is defined as:
\begin{equation}
  \label{eq_smooth}
   S(\dataobs)=\max_{\dataobs' \in \mathcal{X}^{n}}\bigg \{ \frac{1}{\frac{1}{LS(\dataobs')} +\gamma \cdot d(\dataobs,\dataobs')} \bigg \},
\end{equation}
where $d$ is the Hamming distance between two data sets.
\end{definition}

\begin{thm}
  \label{thm_gamma_smooth}
  For any $\dataobs, \dataobs'\in\{0,1\}^n$, and $\gamma \in \mathbb{R}^{+}$
  if  $\adj{\dataobs}{\dataobs'}$ then $\frac{1}{S(\dataobs)} - \frac{1}{S(\dataobs')} \leq \gamma$.
\end{thm}
again, the interested reader will find the proof in Appendix.
We are now ready to describe the mechanism $\hexpmech$ in Algorithm \ref{mech:sexpmech}. $\hexpmech(\dataobs)$ outputs a candidate $r \in \candidateset$ with probability
\[
  \frac {\exp\big(\frac{-\epsilon\cdot \ux{r}}{4\cdot S(\dataobs)}\big)}
  {\sum_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot \ux{r'}}{4\cdot S(\dataobs)})},
\]
where $S(\dataobs)$ is the 1-smooth sensitivity of  $\hellinger(\bysinfer{\dataobs},\cdot)$. {\color{red}{why do we choose 1???}}
  \begin{algorithm}
  \caption{$\hexpmech$}
  \label{mech:sexpmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$,  prior: $\dirichlet{\valpha}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'} = \bys{\valpha}{\dataobs}$.
  \STATE \quad {\bf Initialize} associative map $pr$ of size $|\candidateset|$.
  \STATE \quad {\bf For} $r$ in $\candidateset$:
  \STATE \quad \quad $pr[r] = 
  {\exp(\frac{\epsilon \cdot \ux{r}}{4 S(\dataobs)})}
/ {\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon \cdot \ux{r'}}{4 \cdot S(\dataobs)})}$
\STATE \quad {\bf sample} in $z$ from probability distribution $pr$.
\STATE {\bf return} $z$
  \end{algorithmic}
\end{algorithm}

\begin{thm}
  Algorithm \ref{mech:sexpmech} is $\epsilon$-d.p.
\end{thm}
% As already mentioned $\hexpmech$ extends straightforwardly to the Dirichlet-multinomial system.
% Indeed, the Hellinger distance between two Dirichlet distribution, unfolds in:
% \[
%   \hellinger(\dirichlet{\valpha_1}, \dirichlet{\valpha_2}) = \sqrt{1 - \frac{\betaf(\frac{\valpha_1 + \valpha_2}{2})}{\sqrt{\betaf(\valpha_1) \betaf(\valpha_2)}}},
% \]
% and the $\candidateset$ will consist of the set of all possible posterior Dirichlet
% distributions candidates. Also, the smooth sensitivity $S(\dataobs)$
% in (\ref{thm_gamma_smooth}) will be computed by letting $\dataobs'$ range
% over all the elements in $\datauni^{n}$ adjacent to $\dataobs$. 
\section{Accuracy Analysis}
